---
title: "Untitled"
output: html_document
date: "2025-07-03"
---
# Import necessary packages
```{r}
install.packages("tidyverse")
install.packages("e1071")
install.packages("rpart")
install.packages("rpart.plot")
install.packages("caTools")
install.packages("cluster")
install.packages("factoextra")
install.packages("stats")
install.packages("ggplot2")
install.packages("caret")
install.packages("caret") 
```

```{r}
library(caret)
library(tidyverse)
library(e1071)
library(rpart)
library(rpart.plot)
library(caTools)
library(cluster)
library(factoextra)
library(stats)
library(ggplot2)
```

# Loading the data set
```{r}
data_set <- read.csv("C:\\Users\\dhana\\Downloads\\Sleep_health_and_lifestyle_dataset (1)(Sleep_health_and_lifestyle_data).csv")
data_set3 <- read.csv("C:\\Users\\dhana\\Downloads\\Sleep_health_and_lifestyle_dataset (1)(Sleep_health_and_lifestyle_data).csv")
print("Done")
```
#displays the first 4 rows of the data frame
```{r}
head(data_set, 4)

```
#displays the last 4 rows of the data frame
```{r}
tail(data_set, 4)
```
#displays the dimensions of the dataset
```{r}
print(dim(data_set))
```
#displays the structure of the dataset
```{r}
print(str(data_set))
```
#displays a statistical summary of each column in the dataset
```{r}
print(summary(data_set))
```
#The given R code filters and displays all male software engineers aged 30 or below from the data_set
```{r}
data_set$Age <- as.numeric(data_set$Age)
data_set$Occupation <- as.character(data_set$Occupation)

Filter_Occupation01 <- dplyr::filter(data_set, 
                                     Occupation == "Software Engineer", 
                                     Age <= 30,
                                     Gender == "Male")



Filter_Occupation01
```
#This code filters the dataset to select female Software Engineers aged 30 or below.
```{r}
Filter_Occupation02 <- dplyr::filter(data_set, Occupation == "Software Engineer", Gender == "Female", Age <= 30)
Filter_Occupation02

```
#This code filters the dataset to select male Software Engineers whose sleep duration is 6.2 hours or less and heart rate is 70 or higher.
```{r}
Filter_Occupation03 <- dplyr::filter(data_set, Occupation == "Software Engineer", Gender == "Male", Sleep.Duration <= 6.2, Heart.Rate >= 70)

Filter_Occupation03
```
#This code filters the dataset to select male Doctors aged 30 or younger with a blood pressure reading of "120/80".
```{r}
Filter_Occupation04 <- dplyr::filter(data_set, Occupation == "Doctor", Gender == "Male", Blood.Pressure == "120/80", Age <= 30)
Filter_Occupation04

```
#This code calculates the total number of missing values in the entire dataset and also counts missing values column-wise, then prints both results.
```{r}
missing_data <- sum(is.na(data_set))

missing_data_colunmwise <- colSums(is.na(data_set))

print(missing_data)
print(missing_data_colunmwise)
```
#command in R shows the internal structure of the data_set 
```{r}
str(data_set)
```
#This code creates a blue boxplot showing the age distribution by occupation, highlighting outliers in red.
```{r}
library(dplyr)
library(ggplot2)

data_set %>%
  ggplot(aes(x = Occupation, y = Age)) +
  geom_boxplot(outlier.color = "red",
               outlier.shape = 16,
               outlier.size = 3,
               fill = "blue") +
  labs(title = "Age Distribution by Occupation",
       x = "Occupation",
       y = "Age of Person") +
  theme_bw()
```
#command in R shows the internal structure of the data_set 
```{r}
str(data_set)
```


```{r}
# Assuming dataset03 is your data frame
new_data_set <- data_set

# Define the columns for which you want to calculate the bounds
columns <- c("Age")  # List the column names as strings

# Define the threshold for outliers
outlier_threshold <- 1.5

# Create empty vectors to store the lower and upper bounds
lower_bounds <- c()
upper_bounds <- c()

# Calculate the IQR and bounds for each column using a for loop
for (column in columns) {
  column_iqr <- IQR(data_set[[column]], na.rm = TRUE)
  column_lower_bound <- quantile(new_data_set[[column]], 0.25, na.rm = TRUE) - outlier_threshold * column_iqr
  column_upper_bound <- quantile(new_data_set[[column]], 0.75, na.rm = TRUE) + outlier_threshold * column_iqr

  lower_bounds <- c(lower_bounds, column_lower_bound)
  upper_bounds <- c(upper_bounds, column_upper_bound)
}

# Remove outliers for each column using a for loop
for (i in 1:length(columns)) {
  column <- columns[i]
  lower_bound <- lower_bounds[i]
  upper_bound <- upper_bounds[i]

  data_set <- new_data_set[new_data_set[[column]] >= lower_bound & new_data_set[[column]] <= upper_bound, ]
}

# Create a combined box plot
boxplot_data <- data_set[, columns, drop = FALSE]
boxplot_data %>%
  ggplot(aes( y = Age)) +
  geom_boxplot(outlier.color = "red", outlier.shape = 16, outlier.size = 3, fill = "blue") +
  labs(title = "Age",
       y = "Age of Person") +
         theme_bw()
```
#command in R shows the internal structure of the data_set 
```{r}
str(data_set)
```
#This code counts and prints the total number of duplicate rows in the dataset.
```{r}
duplicate_data <- sum(duplicated(data_set))

print(duplicate_data)
```
#command in R shows the internal structure of the data_set 
```{r}
str(data_set)
```
#This code converts the categorical columns BMI.Category, Blood.Pressure, and Sleep.Disorder into numeric codes by first turning them into factors and then to numbers.
```{r}
data_set$BMI.Category <- as.numeric(factor(data_set$BMI.Category))
data_set$Blood.Pressure <- as.numeric(factor(data_set$Blood.Pressure))
data_set$Sleep.Disorder <- as.numeric(factor(data_set$Sleep.Disorder))
```
#command in R shows the internal structure of the data_set 
```{r}
str(data_set)
```
#This code creates a new dataset new_data_set by removing the Person.ID column from data_set.
```{r}
new_data_set <- subset(data_set, select = -Person.ID)
```
#This code displays the first 3 rows of the new_data_set
```{r}
head(new_data_set, 3)
```
#This code adds a scaled version of the Heart.Rate column to the dataset and displays the first four rows
```{r}
new_data_set <- new_data_set %>%
  mutate(Heart_rate_Scale = scale(new_data_set$Heart.Rate))
head(new_data_set, 4)
```
#This code adds a scaled version of the Daily.Steps column to the dataset and displays the first four rows
```{r}
new_data_set <- new_data_set %>%
  mutate(Daily_Steps_Scale = scale(new_data_set$Daily.Steps))
head(new_data_set, 4)
```
#This code adds a scaled version of the Physical.Activity.Level column to the dataset and displays the first four rows.
```{r}
new_data_set <- new_data_set %>%
  mutate(Physical_Activity_Level_scale = scale(new_data_set$Physical.Activity.Level))
head(new_data_set, 4)
```
#This code plots a scatter plot of scaled heart rate versus scaled daily steps by occupation with a linear trend line.
```{r}
new_data_set %>%
  ggplot(aes(x = Heart_rate_Scale, y = Daily_Steps_Scale, color = Occupation)) +
  geom_point(size = 3, alpha = 0.4) +
  geom_smooth(method = "lm", se = F, color = "black") +
  labs(title = "Scatter plot of Heart rate Vs Daily Steps By Occupation",
                x = "Heart Rate",
                y = "Sleep Duration") +
                  theme_minimal()
```
#This code visualizes the relationship between heart rate and physical activity level by gender using a scatter plot with a linear trend line.
```{r}
new_data_set %>%
  ggplot(aes(x = Heart_rate_Scale, y = Physical_Activity_Level_scale, color = Gender)) +
  geom_point(size = 3, alpha = 0.4) +
  geom_smooth(method = "lm", se = F, color = "black") +
  labs(title = "Scatter Plot of Heart Rate Vs Physical Activity Level of Gender",
                x = "Heart Rate",
                y = "Physical Activity Level") +
                  theme_minimal()
```
#This code visualizes how sleep duration relates to sleep quality across occupations using a scatter plot with a trend line
```{r}
new_data_set %>%
  ggplot(aes(x = Sleep.Duration, y = Quality.of.Sleep, color = Occupation)) +
  geom_point(size = 3, alpha = 0.4) +
  geom_smooth(method = "lm", se = F, color = "black") +
  labs(title = "Scatter Plot of Sleep Duration Vs Quality of Sleep",
                x = "Sleep Duration",
                y = "Quality of Sleep") +
                  theme_minimal()
```
#This code visualizes the relationship between sleep quality and stress level across occupations using a color-coded scatter plot
```{r}
new_data_set$Occupation <- as.factor(new_data_set$Occupation)

new_data_set %>%
  ggplot(aes(x = Quality.of.Sleep, y = Stress.Level, col = Occupation)) +
  geom_point(size = 3, alpha = 0.4) +
  scale_color_brewer(palette = "Set3") +
  labs(title = "Scatter Plot of Quality of Sleep Vs Stress Level",
                x = "Quality of Sleep",
                y = "Stress Level") +
                  theme_minimal()
```
#This code creates a scatter plot of age versus sleep duration, colored and faceted by gender, with a linear trend line and a clean black-and-white theme.
```{r}
new_data_set$Gender <- as.factor(new_data_set$Gender)
ggplot(new_data_set, aes(x = Age, y = Sleep.Duration, color = Gender)) +
  geom_point(size = 3, alpha = 0.4) +
    geom_smooth(method = "lm", se = F, color = "black") +
scale_color_brewer(palette = "Set1") +
facet_wrap(~Gender) +
labs(title = "Scatter Plot of Age vs Sleep Duration by Gender",
             x = "Age",
             y = "Sleep Duration") +
theme_bw()
```
#This code creates a scatter plot showing the distribution of age across different occupations, colored by gender, with a minimal theme applied.
```{r}
new_data_set$Gender <- as.factor(new_data_set$Gender)
ggplot(new_data_set, aes(x = Age, y = Occupation, color = Gender)) +
geom_point(size = 3, alpha = 0.4) +
  #geom_smooth(method = "lm", se = F, color = "black") +
scale_color_brewer(palette = "Set1") +
labs(title = "Scatter Plot of Age vs Occupation by Gender",
              x = "Age",
              y = "Occupation") +
   theme_minimal()
```
#This code creates a scatter plot displaying the relationship between age and BMI category, colored by gender, with a minimal theme.
```{r}
ggplot(new_data_set, aes(x = Age, y = BMI.Category, color = Gender)) +
geom_point(size = 3, alpha = 0.4) +
scale_color_brewer(palette = "Set1") +
labs(title = "Scatter Plot of Age vs BMI Category by Gender",
              x = "Age",
              y = "BMI Category") +
  theme_minimal()
```
#This code creates a dodged bar chart showing the count of people in each age group, separated by gender, with colorful fills and a minimal theme.
```{r}
ggplot(new_data_set, aes(x = factor(Age), fill = Gender)) +
geom_bar(position = "dodge") +
scale_fill_brewer(palette = "Set3") +
labs(title = "Number of People in Each Age Group by Gender",
             x = "Age",
             y = "Count") +
theme_minimal()
```
#This code visualizes the gender distribution across occupations using a dodged bar chart with angled labels for clarity.
```{r}
ggplot(new_data_set, aes(x = Occupation, fill = Gender)) +
geom_bar(position = "dodge") +
scale_fill_brewer(palette = "Set3") +
labs(title = "Number of People in Each Occupation by Gender",
            x = "Occupation",
            y = "Count") +
 theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#This code creates a dodged bar chart showing the distribution of sleep duration by gender, using colorful fills and a minimal theme.
```{r}
ggplot(new_data_set, aes(x = factor(Sleep.Duration), fill = Gender)) +
geom_bar(position = "dodge") +
scale_fill_brewer(palette = "Set3") +
labs(title = "Distribution of Sleep Duration by Gender",
             x = "Sleep Duration (hours)",
             y = "Count") +
theme_minimal()
```
#This code creates a dodged bar chart displaying the distribution of sleep quality by gender with colorful fills and a clean minimal theme
```{r}
ggplot(new_data_set, aes(x = factor(Quality.of.Sleep), fill = Gender)) +
  geom_bar(position = "dodge") +
scale_fill_brewer(palette = "Set3") +
labs(title = "Distribution of Quality of Sleep by Gender",
              x = "Quality of Sleep",
              y = "Count") +
                theme_minimal()
```
#This code creates a dodged bar chart showing the distribution of physical activity levels by gender, using vibrant colors and a minimal theme.
```{r}
ggplot(new_data_set, aes(x = factor(Physical.Activity.Level), fill = Gender)) +
  geom_bar(position = "dodge") +
scale_fill_brewer(palette = "Set3") +
labs(title = "Distribution of Physical Activity Level by Gender",
             x = "Physical Activity Level",
             y ="Count") +
theme_minimal()
```
#This code visualizes stress level distribution across occupations using a color-coded dodged bar chart.
```{r}
ggplot(new_data_set, aes(x = factor(Stress.Level), fill = Occupation)) +
  geom_bar(position = "dodge") +
scale_fill_brewer(palette = "Set3") +
labs(title = "Distribution of Stress Level by Gender",
              x = "Stress Level",
              y = "Count") +
                theme_minimal()
```
#This code creates a dodged bar chart showing the count of people in each BMI category by gender, using colorful fills and a minimal theme.
```{r}
ggplot(new_data_set, aes(x = BMI.Category, fill = Gender)) +
geom_bar(position = "dodge") +
scale_fill_brewer(palette = "Set3") +
labs(title = "Number of People in Each BMI Category",
              x = "BMI Category",
              y = "Count") +
                theme_minimal()
```
#This code visualizes blood pressure distribution across occupations using a color-coded dodged bar chart.
```{r}
ggplot(new_data_set, aes(x = factor(Blood.Pressure), fill = Occupation)) +
   geom_bar(position = "dodge") +
scale_fill_brewer(palette = "Set3") +
labs(title = "Distribution of Blood Pressure by Gender",
              x = "Blood Pressure",
              y = "Count") +
                theme_minimal()
```
#This code creates a dodged bar chart showing the distribution of heart rate values by gender, using colorful fills and a minimal theme.
```{r}
ggplot(new_data_set, aes(x = factor(Heart.Rate), fill = Gender)) +
  geom_bar(position = "dodge") +
scale_fill_brewer(palette = "Set3") +
labs(title = "Distribution of Heart Rate by Gender",
              x = "Heart Rate",
              y = "Count") +
                 theme_minimal()
```
#This code visualizes daily steps distribution across occupations using a color-coded dodged bar chart.
```{r}
ggplot(new_data_set, aes(x = factor(Daily.Steps), fill = Occupation)) +
  geom_bar(position = "dodge") +
scale_fill_brewer(palette = "Set3") +
labs(title = "Distribution of Daily Steps by Gender",
              x = "Daily Steps",
              y = "Count") +
                theme_minimal()
```
#This code creates a dodged bar chart showing the distribution of sleep disorders by gender, using colorful fills and a minimal theme.
```{r}
ggplot(new_data_set, aes(x = factor(Sleep.Disorder), fill = Gender)) +
  geom_bar(position = "dodge") +
scale_fill_brewer(palette = "Set3") +
labs(title = "Distribution of Sleep Disorder by Gender",
              x = "Sleep Disorder",
              y = "Count") +
                theme_minimal()
```
#This code creates a histogram showing the distribution of heart rate values for different occupations, with semi-transparent fills and a minimal theme.
```{r}
new_data_set %>%
  ggplot(aes(x = Heart.Rate, fill = as.factor(Occupation))) +
  geom_histogram(color = '#e9ecef', alpha = 0.6, position = 'identity', bins = 30) +
  labs(title = "Heart Rate By Occupation",
                x = "Heart Rate",
                y = "count") +
                  theme_minimal()
```
#This code creates a box plot to compare the age distribution between genders, with a minimal theme applied.
```{r}
ggplot(new_data_set, aes(x = Gender, y = Age, fill = Gender)) +
  geom_boxplot() +
labs(title = "Box Plot of Age by Gender",
              x = "Gender",
              y = "Age") +
                theme_minimal()
```
#This code creates a box plot showing the age distribution across occupations, with fill colors representing gender, using a minimal theme.
```{r}
ggplot(new_data_set, aes(x = Occupation, y = Age, fill = Gender)) +
  geom_boxplot() +
labs(title = "Box Plot of Occupation by Age",
             x = "Occupation",
             y = "Age") +
               theme_minimal()
```
#This code creates a box plot showing stress level distribution by age, filled by gender, with a minimal theme applied.
```{r}
ggplot(new_data_set, aes(x = Age, y = Stress.Level, fill = Gender)) +
  geom_boxplot() +
labs(title = "Box Plot of Stress Level by Age",
              x = "Age",
              y = "Stress Level") +
                theme_minimal()
```
#This code visualizes age differences across BMI categories by gender using a box plot.
```{r}
ggplot(new_data_set, aes(x = BMI.Category, y = Age, fill = Gender)) +
  geom_boxplot() +
labs(title = "Box Plot of BMI Category by Age",
              x = "BMI Category",
              y = "BMI Category") +
                theme_minimal()
```
#This code visualizes how daily steps vary by age and gender using a box plot.
```{r}
ggplot(new_data_set, aes(x = Age, y = Daily.Steps, fill = Gender)) +
  geom_boxplot() +
labs(title = "Box Plot of Daily Steps by Age",
              x = "Daily Steps",
              y = "Daily Stepsl") +
                theme_minimal()
```
#This code visualizes age distribution across occupations using a density plot with a clear legend below.
```{r}
ggplot(new_data_set, aes(x = Age, fill = Occupation)) +
  geom_density(alpha = 0.5) +
labs(title = "Density Plot of Age by Occupation",
              x = "Age",
              y = "Density") +
                theme_minimal() +
theme(legend.position = "bottom", legend.title = element_blank()) +
guides(fill = guide_legend(nrow = 2))
```
#This code creates a density plot showing the distribution of BMI categories by occupation with semi-transparent fills and a minimal theme.
```{r}
ggplot(new_data_set, aes(x = `BMI.Category`, fill = `Occupation`)) +
   geom_density(alpha = 0.5) +
labs(title = "Density Plot of BMI Category by Occupation ",
              x = "BMI Category",
              y ="Density") +
theme_minimal()
```
#This code creates a density plot visualizing the distribution of daily steps across occupations, with transparent fills, a minimal theme, and a legend positioned below the plot.
```{r}
ggplot(new_data_set, aes(x = Daily.Steps, fill = Occupation)) +
geom_density(alpha = 0.5) +
labs(title = "Density Plot of Daily Steps by Occupation",
              x = "Daily Steps",
              y = "Density") +
                theme_minimal() +
theme(legend.position = "bottom", legend.title = element_blank()) +
guides(fill = guide_legend(nrow = 2))
```
#This code visualizes the distribution of sleep disorders across occupations using a density plot with a clear legend below.
```{r}
ggplot(new_data_set, aes(x = Sleep.Disorder, fill = Occupation)) +
  geom_density(alpha = 0.5) +
labs(title = "Density Plot of Daily Steps by Occupation",
              x = "Sleep Disorder",
              y = "Density")+
theme_minimal() +
theme(legend.position = "bottom", legend.title = element_blank())
```
#This code calculates the number of records for each gender in the dataset and stores the results in gender_counts`.
```{r}
gender_counts <- new_data_set %>% group_by(Gender) %>% summarise(Count = n())
```
#This code creates a pie chart showing the proportion of males and females in the dataset, using colorful fills and a clean, legend-only theme.
```{r}
ggplot(gender_counts, aes(x = "", y = Count, fill = Gender)) +
  geom_bar(width = 1, stat = "identity") +
    coord_polar(theta = "y") +
      scale_fill_brewer(palette = "Set3") +
        labs(title = "Proportion of Males and Females", x = "", y = "") +
          theme_void() +
theme(legend.title = element_blank())
```
#This code calculates the number of records for each occupation in the dataset and saves the results as occupation_counts`.
```{r}
occupation_counts <- new_data_set %>% group_by(Occupation) %>% summarise(Count = n())
```
#This code visualizes the occupation distribution in the dataset as a colorful pie chart.
```{r}
ggplot(occupation_counts, aes(x = "", y = Count, fill = Occupation)) +
  geom_bar(width = 1, stat = "identity") + coord_polar(theta = "y") +
scale_fill_brewer(palette = "Set3") +
labs(title = "Distribution of Occupations", x = "", y = "") +
theme_void() +
theme(legend.title = element_blank())
```
#This code calculates the count of records for each BMI category in the dataset and stores the results in bmi_counts`.
```{r}
bmi_counts <- new_data_set %>% group_by(BMI.Category) %>% summarise(Count = n())
```
#This code creates a colorful pie chart showing the distribution of BMI categories in the dataset, using a clean theme without axis labels.
```{r}
# Ensure BMI.Category is categorical
bmi_counts$BMI.Category <- as.factor(bmi_counts$BMI.Category)

# Create pie chart
ggplot(bmi_counts, aes(x = "", y = Count, fill = BMI.Category)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "Distribution of BMI Categories", x = "", y = "") +
  theme_void() +
  theme(legend.title = element_blank())

```
#The code converts `Gender` and `Occupation` columns from categorical to numeric format using factor encoding.
```{r}
data_set$Gender <- as.numeric(factor(data_set$Gender))
data_set$Occupation <- as.numeric(factor(data_set$Occupation))
```
#command in R shows the internal structure of the data_set 
```{r}
str(data_set)
```
#The command displays the count of each category in the `Sleep.Disorder` column of the dataset.
```{r}
table(data_set$Sleep.Disorder)
```
#The code applies k-means clustering with 3 centers on the first 11 columns of `data_set`, using 20 random starts for better accuracy.
```{r}
set.seed(460)
result <- kmeans(data_set[,1:11], center = 3, nstart = 20)

result
```
#The command creates a cross-tabulation showing how the k-means cluster assignments relate to the Sleep.Disorder` categories in the dataset.
```{r}
table(result$cluster, data_set$Sleep.Disorder)
```
#The code cleans the numeric data by removing missing and infinite values, applies k-means clustering with 3 centers, and visualizes the resulting clusters using a cluster plot.
```{r}
library(cluster)

# Clean numeric data (assuming 1:11 are numeric)
clean_data <- data_set[, 1:11] %>%
  dplyr::select(where(is.numeric)) %>%
  na.omit() %>%
  filter_all(all_vars(!is.infinite(.)))

# Run k-means
set.seed(460)
result <- kmeans(clean_data, centers = 3, nstart = 20)

# Plot clusters
clusplot(clean_data, result$cluster, color = TRUE, shade = TRUE, labels = 0, lines = 0)

```
#The code creates a character vector `tot.withinss` of length 10 to store values, likely for total within-cluster sum of squares from k-means clustering.
```{r}
tot.withinss <- vector(mode = "character", length = 10)
```
#The loop stores total within-cluster sum of squares for 1 to 10 clusters in `tot.withinss`.
```{r}
for (i in 1:10) {
  result <- kmeans(data_set[,1:11], center = i, nstart = 20)
  tot.withinss[i] <- result$tot.withinss
}

plot(1:10, tot.withinss, type = "b", pch = 20)
```
#command in R shows the internal structure of the data_set 
```{r}
str(data_set)
```
#The command shows the frequency of each category in the `Sleep.Disorder` column.
```{r}
table(data_set$Sleep.Disorder)
```
#The code converts the `Sleep.Disorder` column into a factor with specified levels 1, 2, and 3.
```{r}
data_set$Sleep.Disorder <- factor(data_set$Sleep.Disorder, levels = c(1, 2, 3))
```
#The code splits the dataset into 80% training and 20% testing data and shows their dimensions.
```{r}
install.packages("caTools")   # Only once
library(caTools)              # Load the package

set.seed(467)
split_data <- sample.split(data_set$Sleep.Disorder, SplitRatio = 0.8)

train_data <- data_set[split_data, ]
test_data  <- data_set[!split_data, ]

dim(train_data)
dim(test_data)

```
#The code trains a decision tree model, makes predictions on test data, and evaluates performance using a confusion matrix.
```{r}
install.packages("rpart")   # Only once
library(rpart)              # Load every time you run your script

# Build the decision tree model
classify <- rpart(
  formula = Sleep.Disorder ~ Gender + Age + Occupation + Sleep.Duration + Quality.of.Sleep +
    Physical.Activity.Level + Stress.Level + BMI.Category + Blood.Pressure + Heart.Rate + Daily.Steps,
  data = train_data,
  parms = list(split = "gini")
)

# Make predictions
model_prediction <- predict(
  classify,
  newdata = test_data[, c("Gender", "Age", "Occupation", "Sleep.Duration", "Quality.of.Sleep",
                          "Physical.Activity.Level", "Stress.Level", "BMI.Category", "Blood.Pressure",
                          "Heart.Rate", "Daily.Steps")],
  type = "class"
)

# View confusion matrix
table(test_data$Sleep.Disorder, model_prediction)

```
#The code installs (if needed) and loads the `rpart.plot` package, then visualizes the decision tree model using `rpart.plot()`.
```{r}
install.packages("rpart.plot")   # Only needed once
library(rpart.plot)              # Load the package

rpart.plot(classify, cex = 1)
```
#The code evaluates the decision tree model by generating a confusion matrix comparing the actual and predicted sleep disorder classes.
```{r}
model_evaluation <- confusionMatrix(test_data$Sleep.Disorder, model_prediction)
```
#The command displays the performance metrics of the model, including accuracy, class-wise prediction results, and other evaluation statistics.
```{r}
model_evaluation
```

